### 2021-09-08 README.Rmd
Add readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

r_citations <- papaja::cite_r(here::here("docs", "manuscript", "r-references.bib"))
```

This repository contains code for the project "Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers." The manuscript is computationally reproducible using this repository. Note that a few manual formatting tweaks to tables and figures were done for the final [manuscript](https://doi.org/10.1101/2021.08.23.21262478).

To reproduce the manuscript:

1.  Clone this repository
2.  Run `renv::restore()` to ensure you have the necessary packages and versions
3.  Knit [manuscript.Rmd](https://github.com/maia-sh/reg-pub-link/blob/master/docs/manuscript/manuscript.Rmd)

This project uses `r r_citations`.